---
title: "174 Project Russia"
author: "Quyen Le"
date: "2023-03-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(forecast) #forecasting
library(UnitCircle) #unit root test
library(astsa) #sarima modeling
library(MASS) #Box Cox transformation


```

## RUSSIA DATA

##Preparing the data

Reading in the data:

```{r}
russia_data <- read.csv("C:/Users/pbsbd/Desktop/174_proj/RUSSIA_INDEX.csv", header = TRUE, sep=',')
russia_data$DATE = as.Date(russia_data$DATE)
colnames(russia_data)[2] <- 'Index'

#training data set - used to build model
training_russia_data <- head(russia_data, -12) 

#testing data set - later used for forecasting
testing_russia_data <- tail(russia_data, n=12)
```

Plotting the time series:

```{r}
#Converting data to time series variable
russia_data_ts <- ts(training_russia_data[2], start=c(2000,01), frequency = 12)

#Plotting the time series

plot(russia_data_ts, main='Russia CPI for All Items from 2000-01-01 to 2021-03-01', 
     xlab='Year', 
     ylab='CPI April 2020 = 100')
abline(h=0)
```

Below is an attempt to stabilize variance through the Box Cox transformation. The resulting transformed data closely resembles a normal distribution, so we will be using the transformed data for modeling.

```{r}
russia_data_ts_trans <- ts(russia_data_ts + 100, start=c(2000,01), frequency = 12) #shifting data up to apply Box Cox
boxcox(lm(russia_data_ts_trans~1)) #log transformation
title(main='Box Cox Plot to Identify Optimal Transformation')


#Box Cox Transformation
bc_russia_data_ts <- log(russia_data_ts_trans)

#Checking if Box Cox was good
plot(bc_russia_data_ts,
     main = 'Box Cox Transformation of Russia CPI Data')
hist(bc_russia_data_ts,
     main = 'Histogram of Box Cox Transformation of Russia CPI Data')
qqnorm(bc_russia_data_ts, pch = 1, frame = FALSE,
       main = 'QQ Plot of Box Cox Transformation of Russia CPI Data')
qqline(bc_russia_data_ts, col = "steelblue", lwd = 2)
```


##Checking for Stationarity of Transformed Data

Checking for Seasonality:

```{r}
russia_data_components <- decompose(russia_data_ts) #of the training dataset
plot(russia_data_components)
```

Isolating detrended and deseasonalized data (which is now stationary)
and removing NA values:

```{r}
#there are some NA values, let's go ahead and remove them
detrend_russia_data <- na.omit(russia_data_components$random) 
plot(detrend_russia_data,
     main='Russia Stationary Transformed Dataset without NA Values')
```

##Analyzing the Detrended Data's ACF and PACF

Analyzing detrended and deseasonlized data's ACF and PACF plots to determine possible model parameters:

```{r}
#ACF
par(mar = c(4,4,3,1) + 0.1)
acf_detrended_russia <- acf(detrend_russia_data, plot = TRUE, 
                type = 'correlation', 
                main = 'Russia Detrended and Deseasonalized ACF Plot',
                lag.max = 80)
acf_detrended_russia

#PACF 
par(mar = c(4,4,3,1) + 0.1)
pacf_detrended_russia <- pacf(detrend_russia_data, plot = TRUE, 
                  main = 'Russia Detrended and Deseasonalized PACF Plot',
                  lag.max = 80)
pacf_detrended_russia
```


Here are our findings for the ARIMA model parameters from the ACF and the PACF:


Non-seasonal part:
AR - PACF plots - p value
AR(1)
AR(2)
AR(3)
AR(4)
AR(5)
AR(8)


MA - ACF plots - q value
AR(1)
AR(3)
AR(4)
AR(5)
AR(6)
AR(9)


Seasonal part: 
AR - PACF plots - P value
AR(0)


MA - ACF plots - Q value
MA(4) -> 12


Iterating through all possible model parameters:

```{r}
library(astsa)
parameter_opt_russia = list(c(1,2,3,4,5,8), #p
                     1, #d
                     c(1,3,4,5,6,9), #q
                     0, #P
                     1, #D
                     4) #Q
parameter_combos_russia <- expand.grid(parameter_opt_russia)
colnames(parameter_combos_russia) <- c('p', 'd', 'q', 'P', 'D', 'Q')

for(x in 1:nrow(parameter_combos_russia)) {
  current_arima <- NULL
  try(current_arima <- arima(bc_russia_data_ts, #transformed
                             order = c(parameter_combos_russia$p[x], 
                                       parameter_combos_russia$d[x], 
                                       parameter_combos_russia$q[x]),
                             seasonal = list(order = c(parameter_combos_russia$P[x], 
                                                       parameter_combos_russia$D[x], 
                                                       parameter_combos_russia$Q[x]),
                                             period = 12),
                                             method = 'ML'))
  if(!is.null(current_arima) && !'try-error' %in% class(current_arima)) {
    parameter_combos_russia$AIC[x] <- AIC(current_arima)
  }
}
```

##Fitting the Models

After sorting the possible models based on AIC value (lowest to highest), I went down the sorted list to find models that were stationary and passed the following tests:

1. Unit Root Test for Stationarity
2. Portmanteau Tests for Autocorrelation
3. Yule- Walker Test for White Noise in Residuals

```{r}
russia_lowest_AIC <- sort(parameter_combos_russia$AIC)[1:10] #4 lowest numbers
which(parameter_combos_russia$AIC == russia_lowest_AIC[1])#1, STATIONARY
which(parameter_combos_russia$AIC == russia_lowest_AIC[2])#20, STATIONARY 
which(parameter_combos_russia$AIC == russia_lowest_AIC[3])#3, STATIONARY, error in forecasting 
which(parameter_combos_russia$AIC == russia_lowest_AIC[4])#27, STATIONARY
```
Testing Model 1:

```{r}
###MODEL 1###
model_1_russia <- arima(russia_data_ts,
     order = c(parameter_combos_russia$p[1], parameter_combos_russia$d[1], parameter_combos_russia$q[1]),
     seasonal = list(order = c(parameter_combos_russia$P[1], parameter_combos_russia$D[1], parameter_combos_russia$Q[1]),
                     period = 12),
                     method = 'ML')

#unit root testing for model 1 (model stationarity)
roots_1_russia <- data.frame(uc.check(pol_ = c(1, model_1_russia$coef[1]), print_output = FALSE, plot_output = FALSE))
roots_1_russia <- rbind(roots_1_russia,
                  uc.check(pol_ = c(1, model_1_russia$coef[2]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_1_russia$coef[3]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_1_russia$coef[4]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_1_russia$coef[5]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_1_russia$coef[6]), print_output = FALSE, plot_output = FALSE))
roots_1_russia #this a stationary model since all AR values are outside of the unit circle

#getting model residuals
residuals_1_russia <- model_1_russia$residuals

#testing model residuals

#portmanteau tests (correlation)
Box.test(residuals_1_russia, lag = 15, fitdf = 6, type = c("Box-Pierce")) #pass
Box.test(residuals_1_russia, lag = 15, fitdf = 6, type = c("Ljung-Box")) #pass
#thus, residuals have an autocorrelation of 0, which is good

#yule walker tests (white noise)
ar.yw(residuals_1_russia, aic = TRUE, order.max = NULL) 
#order selected = 0, so none of the AR coefficients are significant, therefore the residuals are white noise
```


Testing Model 20:

```{r}
###MODEL 20###
model_20_russia <- arima(russia_data_ts,
     order = c(parameter_combos_russia$p[20], parameter_combos_russia$d[20], parameter_combos_russia$q[20]),
     seasonal = list(order = c(parameter_combos_russia$P[20], parameter_combos_russia$D[20], parameter_combos_russia$Q[20]),
                     period = 12),
                     method = 'ML')

#unit root testing for model 20 (model stationarity)
roots_20_russia <- data.frame(uc.check(pol_ = c(1, model_20_russia$coef[1]), print_output = FALSE, plot_output = FALSE))
roots_20_russia <- rbind(roots_20_russia,
                  uc.check(pol_ = c(1, model_20_russia$coef[2]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_20_russia$coef[3]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_20_russia$coef[4]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_20_russia$coef[5]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_20_russia$coef[6]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_20_russia$coef[7]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_20_russia$coef[8]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_20_russia$coef[9]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_20_russia$coef[10]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_20_russia$coef[11]), print_output = FALSE, plot_output = FALSE))
roots_20_russia #this a stationary model since all AR values are outside of the unit circle

#getting model residuals
residuals_20_russia <- model_20_russia$residuals

#testing model residuals

#portmanteau tests (correlation)
Box.test(residuals_20_russia, lag = 15, fitdf = 11, type = c("Box-Pierce")) #pass
Box.test(residuals_20_russia, lag = 15, fitdf = 11, type = c("Ljung-Box")) #pass
#thus, residuals have an autocorrelation of 0, which is good

#yule walker tests (white noise)
ar.yw(residuals_20_russia, aic = TRUE, order.max = NULL) 
#order selected = 0, so none of the AR coefficients are significant, therefore the residuals are white noise
```

Testing Model 3:

```{r}
###MODEL 3###
model_3_russia <- arima(russia_data_ts,
     order = c(parameter_combos_russia$p[3], parameter_combos_russia$d[3], parameter_combos_russia$q[3]),
     seasonal = list(order = c(parameter_combos_russia$P[3], parameter_combos_russia$D[3], parameter_combos_russia$Q[3]),
                     period = 12),
                     method = 'ML')

#unit root testing for model 3 (model stationarity)
roots_3_russia <- data.frame(uc.check(pol_ = c(1, model_3_russia$coef[1]), print_output = FALSE, plot_output = FALSE))
roots_3_russia <- rbind(roots_3_russia,
                  uc.check(pol_ = c(1, model_3_russia$coef[2]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_3_russia$coef[3]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_3_russia$coef[4]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_3_russia$coef[5]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_3_russia$coef[6]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_3_russia$coef[7]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_3_russia$coef[8]), print_output = FALSE, plot_output = FALSE))
roots_3_russia #this a stationary model since all AR values are outside of the unit circle

#getting model residuals
residuals_3_russia <- model_3_russia$residuals

#testing model residuals

#portmanteau tests (correlation)
Box.test(residuals_3_russia, lag = 15, fitdf = 8, type = c("Box-Pierce")) #pass
Box.test(residuals_3_russia, lag = 15, fitdf = 8, type = c("Ljung-Box")) #pass
#thus, residuals have an autocorrelation of 0, which is good

#yule walker tests (white noise)
ar.yw(residuals_3_russia, aic = TRUE, order.max = NULL) 
#order selected = 0, so none of the AR coefficients are significant, therefore the residuals are white noise
```

##Forecasting the Test Data

Models 1, 20, 3 are ready for forecasting.

Forecasting Model 1:

```{r}
### MODEL 1 ###
russia_data_ts_TEST <- ts(testing_russia_data[2], start=c(2021,04), frequency = 12)
forecast_1_russia <- sarima.for(russia_data_ts, n.ahead = 12,
                         p = 1, d = 1, q = 1, P = 0, D = 1, Q = 4, S = 12,
                         main = "Forecasting of Russia Testing Data, Model 1")
lines(russia_data_ts_TEST, col = 'blue', type = 'l')
legend(x = 'topleft',
       legend = c('Actual', 'Forecasted'),
       lty = c(1, 1),
       col = c('blue', 'red'))

#checking accuracy of model
forecast_1_russia_mse <- mean((russia_data_ts_TEST - forecast_1_russia$pred)**2)
forecast_1_russia_mse #68754.09
```

Forecasting Model 20:

```{r}
### MODEL 20 ###
russia_data_ts_TEST <- ts(testing_russia_data[2], start=c(2021,04), frequency = 12)
forecast_20_russia <- sarima.for(russia_data_ts, n.ahead = 12,
                         p = 2, d = 1, q = 5, P = 0, D = 1, Q = 4, S = 12,
                         main = "Forecasting of Russia Testing Data, Model 20")
lines(russia_data_ts_TEST, col = 'blue', type = 'l')
legend(x = 'topleft',
       legend = c('Actual', 'Forecasted'),
       lty = c(1, 1),
       col = c('blue', 'red'))

#checking accuracy of model
forecast_20_russia_mse <- mean((russia_data_ts_TEST - forecast_20_russia$pred)**2)
forecast_20_russia_mse #68193.34
```

Forecasting Model 3:

```{r}
### MODEL 3 ###
russia_data_ts_TEST <- ts(testing_russia_data[2], start=c(2021,04), frequency = 12)
forecast_3_russia <- sarima.for(russia_data_ts, n.ahead = 12,
                         p = 3, d = 1, q = 1, P = 0, D = 1, Q = 4, S = 12,
                         main = "Forecasting of Russia Testing Data, Model 3")
lines(russia_data_ts_TEST, col = 'blue', type = 'l')
legend(x = 'topleft',
       legend = c('Actual', 'Forecasted'),
       lty = c(1, 1),
       col = c('blue', 'red'))

#checking accuracy of model
forecast_3_russia_mse <- mean((russia_data_ts_TEST - forecast_3_russia$pred)**2)
forecast_3_russia_mse #68946.75
```

model 20 is the best model to use to fit the Russia CPI data.
