---
title: "174 Project UK"
author: "Quyen Le"
date: "2023-03-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(forecast) #forecasting
library(UnitCircle) #unit root test
library(astsa) #sarima modeling
library(MASS) #Box Cox transformation

```

## UK DATA

##Preparing the data

Reading in the data:

```{r}
uk_data <- read.csv("C:/Users/pbsbd/Desktop/174_proj/UK_INDEX.csv", header = TRUE, sep=',')
uk_data$DATE = as.Date(uk_data$DATE)
colnames(uk_data)[2] <- 'Index'

#training data set - used to build model
training_uk_data <- head(uk_data, -12) 

#testing data set - later used for forecasting
testing_uk_data <- tail(uk_data, n=12)

```

Plotting the time series:

```{r}
#Converting data to time series variable
uk_data_ts <- ts(training_uk_data[2], start=c(2000,01), frequency = 12)

#Plotting the time series

plot(uk_data_ts, main='UK Consumer Price Index for All Items from 2000-01-01 to 2021-03-01', 
     xlab='Year', 
     ylab='CPI April 2020 = 100')
abline(h=0)
```
Below is an attempt to stabilize variance through the Box Cox
transformation. The resulting transformed data closely resembles a
normal distribution, so we will be using the transformed data for
modeling.

```{r}
boxcox(lm(uk_data_ts~1))
title(main='Box Cox Plot to Identify Optimal Transformation')

#Box Cox Transformation
BoxCox.lambda(uk_data_ts, lower = -2, upper = 2)
bc_uk_data_ts <- 1/((uk_data_ts)**(1/2)) #inverse square root transformation

#Checking normality of data
plot(bc_uk_data_ts,
     main = 'Box Cox Transformation of uk CPI Data')
hist(bc_uk_data_ts,
     main = 'Histogram of Box Cox Transformation of UK CPI Data')
qqnorm(bc_uk_data_ts, pch = 1, frame = FALSE,
       main = 'QQ Plot of Box Cox Transformation of UK CPI Data')
qqline(bc_uk_data_ts, col = "blue", lwd = 2)
```

##Checking for Stationarity of Transformed Data

Checking for Seasonality:
```{r}
uk_data_components <- decompose(bc_uk_data_ts) 
plot(uk_data_components)
```
Isolating detrended and deseasonalized data (which is now stationary)
and removing NA values:

```{r}
#there are some NA values, let's go ahead and remove them
detrend_uk_data <- na.omit(uk_data_components$random) 
plot(detrend_uk_data,
     main='UK Stationary Transformed Dataset without NA Values')
```

##Analyzing the Detrended Data's ACF and PACF

Analyzing detrended and deseasonlized data's ACF and PACF plots to
determine possible model parameters:

```{r}
#acf of the detrended data
par(mar = c(4,4,3,1) + 0.1)
acf_detrended_uk <- acf(detrend_uk_data, plot = TRUE, 
                type = 'correlation', 
                main = 'UK DIFFERENCED ACF PLOT',
                lag.max = 80)
acf_detrended_uk

#pacf of the detrended data
par(mar = c(4,4,3,1) + 0.1)
pacf_detrended_uk <- pacf(detrend_uk_data, plot = TRUE, 
                  main = 'UK DIFFERENCED PACF PLOT',
                  lag.max = 80)
pacf_detrended_uk
```

Here are our findings for the ARIMA model parameters from the ACF and the PACF:


Non-seasonal part:
AR - PACF plots - p value
AR(1)
AR(2)
AR(3)
AR(4)


MA - ACF plots - q value
AR(1)
AR(2)
AR(4)
AR(5)
AR(6)
AR(7)


Seasonal part: 
AR - PACF plots - P value
AR(2) -> 12


MA - ACF plots - Q value
MA(4) -> 12

Iterating through all possible model parameters:

```{r}
library(astsa)
parameter_opt_uk = list(c(1,2,3,4), #p
                     1, #d
                     c(1,2,4,5,6,7), #q
                     c(2), #P
                     1, #D
                     c(4)) #Q
parameter_combos_uk <- expand.grid(parameter_opt_uk)
colnames(parameter_combos_uk) <- c('p', 'd', 'q', 'P', 'D', 'Q')

for(x in 1:nrow(parameter_combos_uk)) {
  current_arima <- NULL
  try(current_arima <- arima(bc_uk_data_ts, #box cox transformed data
                             order = c(parameter_combos_uk$p[x], parameter_combos_uk$d[x], parameter_combos_uk$q[x]),
                             seasonal = list(order = c(parameter_combos_uk$P[x], parameter_combos_uk$D[x], parameter_combos_uk$Q[x]),
                                             period = 12),
                                             method = 'ML'))
  if(!is.null(current_arima) && !'try-error' %in% class(current_arima)) {
    parameter_combos_uk$AIC[x] <- AIC(current_arima)
  }
}
```

##Fitting the Models

After sorting the possible models based on AIC value (lowest to highest), I went down the sorted list to find models that were stationary and passed the following tests:

1. Unit Root Test for Stationarity
2. Portmanteau Tests for Autocorrelation
3. Yule- Walker Test for White Noise in Residuals

```{r}
uk_lowest_AIC <- sort(parameter_combos_uk$AIC)[1:6]
which(parameter_combos_uk$AIC == uk_lowest_AIC[1]) #11, not stationary
which(parameter_combos_uk$AIC == uk_lowest_AIC[2]) #15, STATIONARY
which(parameter_combos_uk$AIC == uk_lowest_AIC[3]) #5, STATIONARY
which(parameter_combos_uk$AIC == uk_lowest_AIC[4]) #6, error
which(parameter_combos_uk$AIC == uk_lowest_AIC[5]) #3, error 
which(parameter_combos_uk$AIC == uk_lowest_AIC[6]) #9, STATIONARY
```

Testing Model 15:

```{r}
###MODEL 15###
model_15_uk <- arima(bc_uk_data_ts,
     order = c(parameter_combos_uk$p[15], parameter_combos_uk$d[15], parameter_combos_uk$q[15]),
     seasonal = list(order = c(parameter_combos_uk$P[15], parameter_combos_uk$D[15], parameter_combos_uk$Q[15]),
                     period = 12),
                     method = 'ML')

#unit root testing for model 15 (model stationarity)

roots_15_uk <- data.frame(uc.check(pol_ = c(1, model_15_uk$coef[1]), print_output = FALSE, plot_output = FALSE))
roots_15_uk <- rbind(roots_15_uk,
                  uc.check(pol_ = c(1, model_15_uk$coef[2]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_15_uk$coef[3]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_15_uk$coef[4]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_15_uk$coef[5]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_15_uk$coef[6]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_15_uk$coef[7]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_15_uk$coef[8]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_15_uk$coef[9]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_15_uk$coef[10]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_15_uk$coef[11]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_15_uk$coef[12]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_15_uk$coef[13]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_15_uk$coef[14]), print_output = FALSE, plot_output = FALSE)) 
roots_15_uk #this is not a stationary model since SARq value is outside of the unit circle

#getting model residuals
residuals_15_uk <- model_15_uk$residuals

#testing model residuals

#portmanteau tests (correlation)
Box.test(residuals_15_uk, lag = 15, fitdf = 14, type = c("Box-Pierce")) #pass
Box.test(residuals_15_uk, lag = 15, fitdf = 14, type = c("Ljung-Box")) #pass
#thus, residuals have an autocorrelation of 0, which is good

#yule walker tests (white noise)
ar.yw(residuals_15_uk, aic = TRUE, order.max = NULL) 
#order selected = 0, so none of the AR coefficients are significant, therefore the residuals are white noise
```

Testing Model 5:

```{r}
###MODEL 5###
model_5_uk <- arima(bc_uk_data_ts,
     order = c(parameter_combos_uk$p[5], parameter_combos_uk$d[5], parameter_combos_uk$q[5]),
     seasonal = list(order = c(parameter_combos_uk$P[5], parameter_combos_uk$D[5], parameter_combos_uk$Q[5]),
                     period = 12),
                     method = 'ML')

#unit root testing for model 5 (model stationarity)

roots_5_uk <- data.frame(uc.check(pol_ = c(1, model_5_uk$coef[1]), print_output = FALSE, plot_output = FALSE))
roots_5_uk <- rbind(roots_5_uk,
                  uc.check(pol_ = c(1, model_5_uk$coef[2]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_5_uk$coef[3]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_5_uk$coef[4]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_5_uk$coef[5]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_5_uk$coef[6]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_5_uk$coef[7]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_5_uk$coef[8]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_5_uk$coef[9]), print_output = FALSE, plot_output = FALSE)) 
roots_5_uk #this is  a stationary model all AR values are outside of the unit circle

#getting model residuals
residuals_5_uk <- model_5_uk$residuals

#testing model residuals

#portmanteau tests (correlation)
Box.test(residuals_5_uk, lag = 15, fitdf = 9, type = c("Box-Pierce")) #pass
Box.test(residuals_5_uk, lag = 15, fitdf = 9, type = c("Ljung-Box")) #pass
#thus, residuals have an autocorrelation of 0, which is good

#yule walker tests (white noise)
ar.yw(residuals_5_uk, aic = TRUE, order.max = NULL) 
#order selected = 0, so none of the AR coefficients are significant, therefore the residuals are white noise
```

Testing Model 9:

```{r}
###MODEL 9###
model_9_uk <- arima(uk_data_ts,
     order = c(parameter_combos_uk$p[9], parameter_combos_uk$d[9], parameter_combos_uk$q[9]),
     seasonal = list(order = c(parameter_combos_uk$P[9], parameter_combos_uk$D[9], parameter_combos_uk$Q[9]),
                     period = 12),
                     method = 'ML')

#unit root testing for model 9 (model stationary)
roots_9_uk <- data.frame(uc.check(pol_ = c(1, model_9_uk$coef[1]), print_output = FALSE, plot_output = FALSE))
roots_9_uk <- rbind(roots_9_uk,
                  uc.check(pol_ = c(1, model_9_uk$coef[2]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_9_uk$coef[3]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_9_uk$coef[4]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_9_uk$coef[5]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_9_uk$coef[6]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_9_uk$coef[7]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_9_uk$coef[8]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_9_uk$coef[9]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_9_uk$coef[10]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_9_uk$coef[11]), print_output = FALSE, plot_output = FALSE))
roots_9_uk #this is stationary model since AR values are outside of the unit circle

#getting model residuals
residuals_9_uk <- model_9_uk$residuals

#testing model residuals

#portmanteau tests (correlation)
Box.test(residuals_9_uk, lag = 15, fitdf = 11, type = c("Box-Pierce")) #pass
Box.test(residuals_9_uk, lag = 15, fitdf = 11, type = c("Ljung-Box")) #pass
#thus, residuals have an autocorrelation of 0, which is good

#yule walker tests (white noise)
ar.yw(residuals_9_uk, aic = TRUE, order.max = NULL) 
#order selected = 0, so none of the AR coefficients are significant, therefore the residuals are white noise
```

##Forecasting the Test Data

Model 15, 5, and 9 are ready for forecasting.

Forecasting Model 15:

```{r}
### MODEL 15 ###
uk_data_ts_TEST <- ts(testing_uk_data[2], start=c(2021,04), frequency = 12)
forecast_15_uk <- sarima.for(uk_data_ts, n.ahead = 12,
                         p = 3, d = 1, q = 5, P = 2, D = 1, Q = 4, S = 12,
                         main = "Forecasting of UK Testing Data, Model 15")
lines(uk_data_ts_TEST, col = 'blue', type = 'l')
legend(x = 'topleft',
       legend = c('Actual', 'Forecasted'),
       lty = c(1, 1),
       col = c('blue', 'red'))

#checking accuracy of model
forecast_15_uk_mse <- mean((uk_data_ts_TEST - forecast_15_uk$pred)**2)
forecast_15_uk_mse #7.41983
```

Forecasting Model 5:

```{r}
### MODEL 5 ###
uk_data_ts_TEST <- ts(testing_uk_data[2], start=c(2021,04), frequency = 12)
forecast_5_uk <- sarima.for(uk_data_ts, n.ahead = 12,
                         p = 1, d = 1, q = 2, P = 2, D = 1, Q = 4, S = 12,
                         main = "Forecasting of UK Testing Data, Model 5")
lines(uk_data_ts_TEST, col = 'blue', type = 'l')
legend(x = 'topleft',
       legend = c('Actual', 'Forecasted'),
       lty = c(1, 1),
       col = c('blue', 'red'))

#checking accuracy of model
forecast_5_uk_mse <- mean((uk_data_ts_TEST - forecast_5_uk$pred)**2)
forecast_5_uk_mse #7.301916
```

Forecasting Model 9:

```{r}
### MODEL 9 ###
uk_data_ts_TEST <- ts(testing_uk_data[2], start=c(2021,04), frequency = 12)
forecast_9_uk <- sarima.for(uk_data_ts, n.ahead = 12,
                         p = 1, d = 1, q = 4, P = 2, D = 1, Q = 4, S = 12,
                         main = "Forecasting of UK Testing Data, Model 9")
lines(uk_data_ts_TEST, col = 'blue', type = 'l')
legend(x = 'topleft',
       legend = c('Actual', 'Forecasted'),
       lty = c(1, 1),
       col = c('blue', 'red'))

#checking accuracy of model
forecast_9_uk_mse <- mean((uk_data_ts_TEST - forecast_9_uk$pred)**2)
forecast_9_uk_mse #7.318678
```

model 5 is the best model to use to fit the UK CPI data.
