---
title: "174 Project China"
author: "Quyen Le"
date: "2023-03-12"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(forecast) #forecasting
library(UnitCircle) #unit root test
library(astsa) #sarima modeling
library(MASS) #Box Cox transformation

```

#CHINA DATA

##Preparing the data

Reading in the data:

```{r}
china_data <- read.csv("C:/Users/pbsbd/Desktop/174_proj/CHINA_INDEX.csv", header = TRUE, sep=',')
china_data$DATE = as.Date(china_data$DATE)
colnames(china_data)[2] <- 'Index'

#training data set - used to build model
training_china_data <- head(china_data, -12)

#testing data set - later used for forecasting
testing_china_data <- tail(china_data, n=12)
```

Plotting the time series:

```{r}
#Converting data to time series variable
china_data_ts <- ts(training_china_data[2], start=c(2000,01), frequency = 12)

#Plotting the time series

plot(china_data_ts, main='China Consumer Price Index for All Items from 2000-01-01 to 2021-03-01', 
     xlab='Year', 
     ylab='CPI April 2020 = 100')
abline(h=0)

```

Below is an attempt to stabilize variance through the Box Cox
transformation. The resulting transformed data closely resembles a
normal distribution, so we will be using the transformed data for
modeling.

```{r}
china_data_ts_trans <- ts(china_data_ts + 75, start=c(2000,01), frequency = 12) #shifting data up to apply Box Cox
boxcox(lm(china_data_ts_trans~1)) #square root transformation
title(main='Box Cox Plot to Identify Optimal Transformation')

#Box Cox Transformation
bc_china_data_ts <- sqrt(china_data_ts_trans)

#Checking normality of data
plot(bc_china_data_ts,
     main = 'Box Cox Transformation of China CPI Data')
hist(bc_china_data_ts,
     main = 'Histogram of Box Cox Transformation of China CPI Data')
qqnorm(bc_china_data_ts, pch = 1, frame = FALSE,
       main = 'QQ Plot of Box Cox Transformation of China CPI Data')
qqline(bc_china_data_ts, col = "blue", lwd = 2)
```


##Checking for Stationarity of Transformed Data

Checking for Seasonality:

```{r}
#using Box Cox transformed data
china_data_components <- decompose(bc_china_data_ts)
plot(china_data_components)
```

Isolating detrended and deseasonalized data (which is now stationary)
and removing NA values:

```{r}
#there are some NA values, let's go ahead and remove them
detrend_china_data <- na.omit(china_data_components$random) 
plot(detrend_china_data,
     main='China Stationary Transformed Dataset without NA Values')
```

##Analyzing the Detrended Data's ACF and PACF

Analyzing detrended and deseasonlized data's ACF and PACF plots to
determine possible model parameters:

```{r}
#ACF
par(mar = c(4,4,3,1) + 0.1)
acf_detrended_china <- acf(detrend_china_data, plot = TRUE, 
                type = 'correlation', 
                main = 'China Detrended and Deseasonalized ACF Plot',
                lag.max = 80)
acf_detrended_china

#PACF 
par(mar = c(4,4,3,1) + 0.1)
pacf_detrended_china <- pacf(detrend_china_data, plot = TRUE, 
                  main = 'China Detrended and Deseasonalized PACF Plot',
                  lag.max = 80)
pacf_detrended_china
```


Here are our findings for the ARIMA model parameters from the ACF and
the PACF:


Non-seasonal part: 
AR - PACF plots - p value 
AR(1) 
AR(3) 
AR(9) 
AR(10)


MA - ACF plots - q value 
AR(1) 
AR(2) 
AR(10)


Seasonal part: 
AR - PACF plots - P value 
AR(1) 
AR(2)


MA - ACF plots - Q value 
MA(1) 
MA(3)


Iterating through all possible model parameters:

```{r}
parameter_opt_china = list(c(1,3,9,10),
                     1,
                     c(1,2,10),
                     c(1,2),
                     1,
                     c(1,3))
#matrix of all possible parameters
parameter_combos_china <- expand.grid(parameter_opt_china)
colnames(parameter_combos_china) <- c('p', 'd', 'q', 'P', 'D', 'Q') #matrix of all possible parameters

for(x in 1:nrow(parameter_combos_china)) {
  current_arima <- NULL
  try(current_arima <- arima(bc_china_data_ts, #box cox transformed data
                             order = c(parameter_combos_china$p[x], 
                                       parameter_combos_china$d[x], 
                                       parameter_combos_china$q[x]),
                             seasonal = list(order = c(parameter_combos_china$P[x], 
                                                       parameter_combos_china$D[x], 
                                                       parameter_combos_china$Q[x]),
                                             period = 12),
                                             method = 'ML'))
  
  if(!is.null(current_arima) && !'try-error' %in% class(current_arima)) {
    parameter_combos_china$AIC[x] <- AIC(current_arima)
  }
}
```


##Fitting the Models

After sorting the possible models based on AIC value (lowest to highest), I went down the sorted list to find models that were stationary and passed the following tests:

1. Unit Root Test for Stationarity
2. Portmanteau Tests for Autocorrelation
3. Yule- Walker Test for White Noise in Residuals

```{r}
china_lowest_AIC <- sort(parameter_combos_china$AIC)[1:12] 

which(parameter_combos_china$AIC == china_lowest_AIC[1]) #42, not stationary
which(parameter_combos_china$AIC == china_lowest_AIC[2]) #44, NA
which(parameter_combos_china$AIC == china_lowest_AIC[3]) #48, model error
which(parameter_combos_china$AIC == china_lowest_AIC[4]) #20, fails portmanteau
which(parameter_combos_china$AIC == china_lowest_AIC[5]) #40, not stationary
which(parameter_combos_china$AIC == china_lowest_AIC[6]) #46, not stationary
which(parameter_combos_china$AIC == china_lowest_AIC[7]) #16, STATIONARY
which(parameter_combos_china$AIC == china_lowest_AIC[8]) #15, STATIONARY
which(parameter_combos_china$AIC == china_lowest_AIC[9]) #36, model error
which(parameter_combos_china$AIC == china_lowest_AIC[10]) #24, NAs produced
which(parameter_combos_china$AIC == china_lowest_AIC[11]) #12, model error
which(parameter_combos_china$AIC == china_lowest_AIC[12]) #26, STATIONARY
```

Testing Model 16:

```{r}
###MODEL 16###
model_16_china <- arima(bc_china_data_ts,
     order = c(parameter_combos_china$p[16], parameter_combos_china$d[16], parameter_combos_china$q[16]),
     seasonal = list(order = c(parameter_combos_china$P[16], parameter_combos_china$D[16], parameter_combos_china$Q[16]),
                     period = 12),
                     method = 'ML')

#unit root testing for model 2 (model stationary)
roots_16_china <- data.frame(uc.check(pol_ = c(1, model_16_china$coef[1]), print_output = FALSE, plot_output = FALSE))
roots_16_china <- rbind(roots_16_china,
                  uc.check(pol_ = c(1, model_16_china$coef[2]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_16_china$coef[3]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_16_china$coef[4]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_16_china$coef[5]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_16_china$coef[6]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_16_china$coef[7]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_16_china$coef[8]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_16_china$coef[9]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_16_china$coef[10]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_16_china$coef[11]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_16_china$coef[12]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_16_china$coef[13]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_16_china$coef[14]), print_output = FALSE, plot_output = FALSE))
roots_16_china #this is a stationary model since AR values are all outside of the unit circle

#getting model residuals
residuals_16_china <- model_16_china$residuals

#testing model residuals

#portmanteau tests (correlation)
Box.test(residuals_16_china, lag = 15, fitdf = 14, type = c("Box-Pierce")) #pass
Box.test(residuals_16_china, lag = 15, fitdf = 14, type = c("Ljung-Box")) #pass
#thus, residuals does not have an autocorrelation of 0, which is bad

#yule walker tests (white noise)
ar.yw(residuals_16_china, aic = TRUE, order.max = NULL) 
#order selected = 0, so none of the AR coefficients are significant, therefore the residuals are white noise
```

Testing Model 15:

```{r}
###MODEL 15###
model_15_china <- arima(bc_china_data_ts,
     order = c(parameter_combos_china$p[15], parameter_combos_china$d[15], parameter_combos_china$q[15]),
     seasonal = list(order = c(parameter_combos_china$P[15], parameter_combos_china$D[15], parameter_combos_china$Q[15]),
                     period = 12),
                     method = 'ML')

#unit root testing for model 15 (model stationary)
roots_15_china <- data.frame(uc.check(pol_ = c(1, model_15_china$coef[1]), print_output = FALSE, plot_output = FALSE))
roots_15_china <- rbind(roots_15_china,
                  uc.check(pol_ = c(1, model_15_china$coef[2]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_15_china$coef[3]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_15_china$coef[4]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_15_china$coef[5]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_15_china$coef[6]), print_output = FALSE, plot_output = FALSE), 
                  uc.check(pol_ = c(1, model_15_china$coef[7]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_15_china$coef[8]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_15_china$coef[9]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_15_china$coef[10]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_15_china$coef[11]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_15_china$coef[12]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_15_china$coef[13]), print_output = FALSE, plot_output = FALSE))
roots_15_china #this is a stationary model since AR values are all outside of the unit circle

#getting model residuals
residuals_15_china <- model_15_china$residuals

#testing model residuals

#portmanteau tests (correlation)
Box.test(residuals_15_china, lag = 15, fitdf = 13, type = c("Box-Pierce")) #pass
Box.test(residuals_15_china, lag = 15, fitdf = 13, type = c("Ljung-Box")) #pass
#thus, residuals have an autocorrelation of 0, which is good

#yule walker tests (white noise)
ar.yw(residuals_15_china, aic = TRUE, order.max = NULL) 
#order selected = 0, so none of the AR coefficients are significant, therefore the residuals are white noise

```

Testing Model 26:

```{r}
model_26_china <- arima(bc_china_data_ts,
     order = c(parameter_combos_china$p[26], parameter_combos_china$d[26], parameter_combos_china$q[26]),
     seasonal = list(order = c(parameter_combos_china$P[26], parameter_combos_china$D[26], parameter_combos_china$Q[26]),
                     period = 12),
                     method = 'ML')

#unit root testing
roots_26_china <- data.frame(uc.check(pol_ = c(1, model_26_china$coef[1]), print_output = FALSE, plot_output = FALSE))
roots_26_china <- rbind(roots_26_china,
                  uc.check(pol_ = c(1, model_26_china$coef[2]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_26_china$coef[3]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_26_china$coef[4]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_26_china$coef[5]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_26_china$coef[6]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_26_china$coef[7]), print_output = FALSE, plot_output = FALSE),
                  uc.check(pol_ = c(1, model_26_china$coef[8]), print_output = FALSE, plot_output = FALSE))
roots_26_china #this is a stationary model since AR values are all outside of the unit circle

#testing model residuals

#model residuals
residuals_26_china <- model_26_china$residuals

#portmanteau tests (correlation)
Box.test(residuals_26_china, lag = 15, fitdf = 8, type = c("Box-Pierce")) #pass
Box.test(residuals_26_china, lag = 15, fitdf = 8, type = c("Ljung-Box")) #pass
#thus, residuals have an autocorrelation of 0, which is good

#yule walker tests (white noise)
ar.yw(residuals_26_china, aic = TRUE, order.max = NULL) 
#order selected = 0, so none of the AR coefficients are significant, therefore the residuals are white noise
```


##Forecasting the Test Data

model 16, 15, and 26 are going to be good for forecasting, so let's go
ahead and forecast the model.

```{r}
### MODEL 16 ###
china_data_ts_TEST <- ts(testing_china_data[2], start=c(2021,04), frequency = 12)
forecast_16_china <- sarima.for(china_data_ts, n.ahead = 12,
                         p = 10, d = 1, q = 1, P = 2, D = 1, Q = 1, S = 12,
                         main = "Forecasting of China Testing Data, Model 16")
lines(china_data_ts_TEST, col = 'blue', type = 'l')
legend(x = 'topleft',
       legend = c('Actual', 'Forecasted'),
       lty = c(1, 1),
       col = c('blue', 'red'))

#checking accuracy of model
forecast_16_china_mse <- mean((china_data_ts_TEST - forecast_16_china$pred)**2)
forecast_16_china_mse #2215
```

```{r}
### MODEL 15 ###
china_data_ts_TEST <- ts(testing_china_data[2], start=c(2021,04), frequency = 12)
forecast_15_china <- sarima.for(china_data_ts, n.ahead = 12,
                         p = 9, d = 1, q = 1, P = 2, D = 1, Q = 1, S = 12,
                         main = "Forecasting of China Testing Data, Model 15")

lines(china_data_ts_TEST, col = 'blue', type = 'l')
legend(x = 'topleft',
       legend = c('Actual', 'Forecasted'),
       lty = c(1, 1),
       col = c('blue', 'red'))

#checking accuracy of model
forecast_15_china_mse <- mean((china_data_ts_TEST - forecast_15_china$pred)**2)
forecast_15_china_mse #1484.361
```

```{r}
### MODEL 26 ###
china_data_ts_TEST <- ts(testing_china_data[2], start=c(2021,04), frequency = 12)
forecast_26_china <- sarima.for(china_data_ts, n.ahead = 12,
                         p = 3, d = 1, q = 1, P = 1, D = 1, Q = 3, S = 12,
                         main = "Forecasting of China Testing Data, Model 26")

lines(china_data_ts_TEST, col = 'blue', type = 'l')
legend(x = 'topleft',
       legend = c('Actual', 'Forecasted'),
       lty = c(1, 1),
       col = c('blue', 'red'))

#checking accuracy of model
forecast_26_china_mse <- mean((china_data_ts_TEST - forecast_26_china$pred)**2)
forecast_26_china_mse #200.2951
```

model 26 is the best model to use to fit the China CPI data.
